msre reactor design

### 2019-09-10 report/msrecore.tex
several added pictures and typewriter font

\documentclass{article}
\usepackage{geometry}
\geometry{paperwidth=170mm,
          paperheight=2500mm,
          top=20mm,
          bottom=20mm
          }
\usepackage{float}

@@

\title{msre core}
\author{}
\date{}

@@

\section{introduction}
the design and dimensions of the molten salt reactor experiment is detailed in several ornl reports. this document aims at providing a comprehensive overview of the core design, dimensions, and materials, for the purpose of creating an accurate cad model of the reactor core. \\

@@

\section{msre core}
an isometric view of the msre core can be seen in figure \ref{3229-fig1}
\begin{figure}[H]
  \centering
  \fbox{\includegraphics[page=6,width=0.8\textwidth,trim={2cm 6cm 3.5cm 4cm},clip]{\mrsarchive/docs/ORNL-TM-3229.pdf}}
  \caption{msre reactor vessel \parencite[figure 1]{ornl-tm-3229}.}
  \label{3229-fig1}
\end{figure}

@@

\section{core vessel}
\subsection{lower and upper vessel plenum}
\textcite[page 14]{ornl-tm-3229} states about the lower vessel plenum:
\enquote{\textit{the lower plenum of the core vessel is formed by a standard 60 in. od asme flanged and dished head}} and \textcite[page 25]{ornl-tm-3229} states about the upper vessel plenum:
\enquote{\textit{the reactor vessel upper plenum is similar to the lower plenum in that it is forme from a standard asme flanged and dished head}}.

\subsection{flow distributor and annulus}
\textcite[page 112 and 114]{ornl-tm-3039} states about the annulus:
\enquote{\textit{the fuel salt entered the flow distributor where it was directed downward around the circumference of the vessel. it flowed in a spiral path through a 1-in. annulus between the vessel wall and the core can for cooling purposes.}}. \enquote{\textit{to prevent possible overheating in an otherwise stagnant region, a small portion of salt entering the reactor was diverted into the region just above the core can support flange in the annulus between the vessel and the core can}}.
this is depicted in figure \ref{0728-p109-p110}. \\

@@

\subsubsection{flow distributor orifices}
\textcite[page 9-10]{ornl-tm-3229} states about the flow distributor orifices:
\enquote{\textit{one characteristic of this type of volute is the variable static pressure around it; therefore, in order to obtain a uniform angular flow distribution, it is necessary to use orifices between the volute and core wall cooling annulus. the orifices are 3/4 in. and occur in stacks of 3. at the head end of the volute the orifice stacks are 5 deg apart. at the tail end of the volute, because of the lower fuel velocity and resulting higher static pressure, the orifice stack spacing is increased to 22 1/2 deg. this orifice distribution was determined from the 1/5 scale model. the orifice holes are drilled at an angle of 30 deg with the tangent in order to maintain a tangential velocity component in the annulus. the resulting high heat transfer coeffficients cool the core vessel wall and the reactor core can}}.

@@

\subsubsection{bypass flow, core support lugs and flange}
\parencite[page 85]{ornl-tm-0728} states about the bypass:
\enquote{\textit{to prevent possible overheating in a region that might otherwise have been stagnant, about 24 gpm of salt entering the reactor is diverted into the region just above the core-can support flange in the annulus between the pressure vessel and the core can. this is accomplished through 18 slots or channels, 0.2 by 0.2 in., cut in the core-can flange. these slots are machined at an angle of $30^\circ$ to promote better mixing in the region. in addition, a by-pass flow of 3-22 gpm of salt will pass through the annular clearances at the core can support ring}}.

@@

\subsection{anti-swirl vanes}
\textcite[page 14]{ornl-tm-3229} states about the anti-swirl vanes:
\enquote{\textit{the anti-swirl vanes consist of 48 plates starting about 2 in. up in the core wall cooling annulus and extending along radial lines into the lower head for about 38\% of the radial distance to the core centerline. they are slightly elevated of the core vessel wall, thus eliminating as much coner area as possible where settled solids could accumulate.}}. \\

@@

\subsection{vessel drain}
\label{sec:drain}
\textcite[page 14-15]{ornl-tm-3229} states about the the two drain lines:
\enquote{\textit{the vessel drain consist of a short section of 1 1/2 in. pipe extending slightly up into the vessel head at the centerline, and having a conical umbrella over it. in addition it incorporates a secondary drain with is a 1/2 in. contentric tube coming up the middel of the primary drain, penetrating through the drain pipe wall just inside the vessel and wrapping around the drain pipe horizontally about $90^\circ$}}.

@@

\subsection{grid support plates}
\parencite[page 20]{ornl-tm-3229} states about the grid support structure:
\enquote{\textit{in flowing through the reactor moderator assembly, fuel first passes through a moderator support structure and then through the moderator core blocks. the support structure consists of two assemblies, the lower of which is a hastelloy-n cross structure ... which is the main support structure for the graphite}}.

@@

\begin{figure}[H]
  \centering
  \centering
  \fbox{\includegraphics[page=165,width=0.8\textwidth,trim={4cm 5.5cm 4cm 2.25cm},clip]{\mrsarchive/docs/ORNL-4676.pdf}}
  \caption{comparison of the bottom positioning stud on the graphite moderator element when initially installed in the msre and after operation \parencite[figure 13.3]{ornl-4676}.}
  \label{4676-fig13-3}
\end{figure}

@@

\subsection{vessel outlet}
\parencite[page 84-85]{ornl-tm-0728} states about the vessel outlet:
\enquote{\textit{the salt leaves the reactor core and flows through the upper head to the 10-in. nozzle opening. it is diverted through a 5-in. opening in the side of the nozzle to flow to the fuel circulating pump. the 10-in. nozzle also serves as an access port and support for the tree control rods and for taking and placing of the four graphite-sample rods in the core matrix}}.

@@

\begin{figure}[H]
  \centering
  \fbox{\includegraphics[page=87,width=0.8\textwidth,trim={3cm 4cm 3.75cm 3cm},clip]{\mrsarchive/docs/ORNL-3708.pdf}}
  \caption{complete control rod thimble assembly \parencite[figure 44]{ornl-3708}.}
  % \label{3708-fig40}
\end{figure}

@@

\subsubsection{removable nozzle plug}
\parencite[page 104-105]{ornl-tm-0728} states about the removable nozzle plug:
\enquote{\textit{the removable nozzle plug is about 20-1/2 in. long and is 9.770 in od at the top and 9.520 in. od at the bottom to provide a taper to assist in freeing the plug for removal. the radial clearance is 1/8 in. at the top and 1/4 in. at the bottom. fuel salt is frozen in the annular space between the plug and the nozzle, to effect a salt seal. the salt maintained below the freezing temperature by cooling the outside of the nozzle extension and the inside of the plug with a flow of cell atmosphere gas}}.
\enquote{\textit{the salt seal prevents the salt from coming in contact with and corroding the ring-joint gas seal on the upper flange}}.
\enquote{\textit{the nozzle plug is hollow and is filled with insulation. the upper end of the plug is welded to the mating flange for the 10-in. closure}}.

@@

\subsubsection{outlet strainer}
\textcite[page 24-26]{ornl-tm-3229} states about the outlet strainer:
\enquote{\textit{after the model was built, a strainer structure ... was added to the reactor outlet design which penetrated down into the upper plenum. its purpose was to catch graphite chips (graphite floats in fuel salt) should they break loose from the moderator. this device was not installed in the model but it would not be expected to incluence the flow significantly}}. \\

@@

\textcite[page 115]{ornl-tm-3039} states about the outlet strainer:
\enquote{\textit{stainers were provided at the reactor outlet to prevent passage of graphite chips larger then 1/16 in. a strainer basket was attached to the lower end of the 10-in. nozzle plug and extended downward into the upper head region of the reactor vessel. the three control rod timbles and the graphite sample assembly passed through the stainer basket. since the five graphite elements were not pinned as were the remaining moderator elements, a cross-shaped extension of the basket essembly projected beneath the basket to provide a hold-down}}. this is depicted in figure \ref{3039-fig5-11}. \\

@@

\section{graphite can}
the msre used a graphite moderator consisting of an assembly of graphite stringer with control rods and removable samples rods in the center. bellow is seen the assembly of the can without the removable rods at the center.
\begin{figure}[H]
  \centering
  \fbox{
    \includegraphics[width=0.4\textwidth]{\mrsarchive/img/core-half.jpg}
    \includegraphics[width=0.393\textwidth]{\mrsarchive/img/core-full.jpg}
  }
  \caption{msre graphite core assembly \parencite{ornl}.}
  \label{core-assembly}
\end{figure}
\textcite[page 112 and 114]{ornl-tm-3039} states about the graphite stringers: \enquote{\textit{the reactor core consisted of 617 full and factional sized graphite elements 2x2 in. cross section and about 67 in. long}}. \enquote{\textit{when not buoyed by being immersed in salt, the vertical graphite elements were supported by a lattice of graphite blocks which in turn are supported by a hastelloy-n grid fastened to the bottom of the core can. the core can was thus free to expand downward relative to its top attachment to the reactor vessel while the graphite was free to expand upward relative to its support at the lower end of the core can. the graphite moderator elements were restrained from floating out of core by hastelloy-n rod through holes in the dowel section at the bottom of each graphite element. in addition a wire passing through the upper graphite elements prevented the upper partion of a broken element from floating out of the core}}.

@@

\subsection{typical graphite stringer}
a typical graphite stringers, along with a removable graphite stringer, is given in figure \ref{3229-fig2}
\begin{figure}[H]
  \centering
  \centering
  \fbox{\includegraphics[page=7,width=0.8\textwidth,trim={3.5cm 6cm 2.5cm 5cm},clip]{\mrsarchive/docs/ORNL-TM-3229.pdf}}
  \caption{typical graphite core block arrangement \parencite[figure 2]{ornl-tm-3229}.}
  \label{3229-fig2}
\end{figure}

@@

\subsubsection{dowel section}
the dowel at the bottom consists of two section dowel (long slim part and stumpy thick part) seen below.
\begin{figure}[H]
  \centering
  \fbox{\includegraphics[page=385,width=0.8\textwidth,trim={2.5cm 6.2cm 3.3cm 4.1cm},clip]{\mrsarchive/docs/ORNL-3708.pdf}}
  \caption{appearance of spalls and cracks in samples of severly cracked grade cgb graphite in (a) spindle end of a machined bar and (b) midplane of an as-received bar. cracks are indicated by arrows or are clearly visible as ragged white lines. \parencite[figure 4]{ornl-3708}.}
  \label{3708-fig4}
\end{figure}

@@

\subsection{hold-down rods and support wires}
\label{sec:hold-down-rods-support-wires}
\parencite[page 84-85]{ornl-tm-0728} states about the hold-down rods and support wires:
\enquote{\textit{a 5/16-in.-diam inor-8 rod passes through a 0.010-in.-wall-thichness bushing placed in a 0.375-in.-diam hole in the dowel section at the bottom of each graphite stringer. these rods also pass through the inor-8 grid-supporting structure and prevent each graphite stringer from floating in the fuel salt. if a graphite stringer were to break in two, the top partion would tend to float away and leave a relatively stagnant pocket of fuel salt which might reach a higher temperature than desired. the effect of this on the reactivity and on the temperatures in the reactor has been studied. to guard against this eventuality, a 1/16-in.-diam inor-8 wire is passed through a 1/8-in.-diam inor-8 insert about 1 in from the top of each graphite stringer, fastening the tops together and to the core can}}.

@@

\subsection{graphite lattice blocks}
\parencite[page 20]{ornl-tm-3229} states about the lattice blocks: \enquote{\textit{resting on the this}(grid support plates)\textit{ is a grid consisting of two layers of rectangular graphite bars, one layer resting on the other and perpendicular to it. the purpose of the graphite assembly is to position and hold the graphite core blocks, and to compensate for a difference in thermal expansion between the hastelloy n and graphite}}.

@@

\subsection{removable graphite stringer}
\parencite[page 84]{ornl-tm-0728} states about the removable graphite stringer:
\enquote{\textit{the five graphite stringers at the center of the core can be removed, although with considerable more difficulty}}.
see figure \ref{0728-fig5-7}.
\enquote{\textit{the five stringers are of a special design (types 7, 60 and 61 on ornl dwgs d-bb-b-40416, 40418 and 40581). they are 2-in. x 2-in. in cross section but are 64-1/2 in. long rather than the 62-1/8 in. of the average stringer. they do not have the dowel section at the bottom and the hole for the hold-down rods and they rest directly on the inor-8 supporting grid rather than on the graphite lattice blocks. the lattice blocks do not extend across the five stringers locations, the opening thus provided through the blocks permitting insertion of a viewing device through the core to permit observation of the lower head should this prove desirable}}.
\enquote{\textit{each of the five stringers is drilled and tapped with a 3/4-in. 6 acme thread on the upper end for an inor-8 lifting knob, or stud, which can engage a snaptite quick-disconnect coupling. they are prevented from floating in the fuel salt by 1/8 in. x 1/2 in. hold-down bars welded to the strainer assembly at a level even with the top of the graphite core. (see ornl dwg e-bb-b-40598)}}.
see the figure \ref{3229-fig2}.

@@

\begin{figure}[H]
  \centering
  \centering
  \fbox{\includegraphics[page=164,width=0.8\textwidth,trim={3.5cm 11.5cm 2cm 3.5cm},clip]{\mrsarchive/docs/ORNL-4676.pdf}}
  \caption{... \parencite[figure 13.2]{ornl-4676}.}
  \label{4676-fig13-2}
\end{figure}

% \subsection{tall graphite stringers}

\subsection{core centering grid or graphite centering bridge}
\parencite[page 84-85]{ornl-tm-0728} states about the core centering grid:
\enquote{\textit{at the top of the graphite blocks a centering bridge holds a row of stringers in position on two diameters at right angles to each other. this bridge helps to prevent shifting of the entire stringer assembly (see ornl dwg. d-bb-b-40424)}}.
\begin{figure}[H]
  \centering
  \centering
  \fbox{\includegraphics[page=29,width=0.8\textwidth,trim={3cm 13.7cm 4cm 4cm},clip]{\mrsarchive/docs/ORNL-4676.pdf}}
  \caption{fish eye view of top of core assembly with centering bridge and broken pieces of graphite \parencite[figure 1.5]{ornl-4676}.}
  \label{4676-fig1-5}
\end{figure}

@@

\subsection{restrictor ring}
\parencite[page 84-85]{ornl-tm-0728} states about the restrictor ring:
\enquote{\textit{to prevent an excessive amount of salt flow in the annuluar space thus created, an inor-8 restrictor ring, 1/2 x 1/2 x 54-1/2 in. id. surrounds the bundle of graphite stringers at the bottom (see ornl dwg. d-bb-b-40427). the stringers are restrained from excessive movement at the top by a graphite retainer ring, 1 x 2 x 53-1/4 in. id. this ring, in turn, is held in place by an inor-8 retainer ring, 3/4 x 3/4 x 53-1/4 in. id (see ornl dwg. d-bb-b-40428 for both rings)}}.

\subsection{core can}
\parencite[page 79]{ornl-tm-0728} states about the core can:
\enquote{\textit{the core can, or shell, is 55-1/2 in. ID and 67-15/16 in. high and was rolled from 1/4-in.-thick INOR-8 plate.}}

\section{control stringer and sample assembly}
the control stringer and sample assembly is further detailed in figure \ref{61246-fig14a}, \ref{61246-fig14b}, and \ref{61246-fig14c}.
\begin{figure}[H]
  \centering
  \fbox{\includegraphics[page=117,width=0.8\textwidth,trim={2.7cm 5.5cm 3.1cm 4.2cm},clip,angle=0.4]{\mrsarchive/docs/ORNL-TM-0728.pdf}}
  \caption{lattice arrangement of control rods and sample rod \parencite[figure 5.7]{ornl-tm-0728}.}
  \label{0728-fig5-7}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \fbox{
    \begin{tabular}{c}
      \includegraphics[page=118,width=0.8\textwidth,trim={11cm 5cm 0cm 4cm},clip,angle=-90]{\mrsarchive/docs/ORNL-TM-0728.pdf} \\ [-0.3cm]
      \includegraphics[page=119,height=0.50\textwidth,trim={0cm 4.4cm 33cm 3cm},clip,angle=-90]{\mrsarchive/docs/ORNL-TM-0728.pdf}
    \end{tabular}
  }
  \caption{graphite-inor-8 sample assembly \parencite[figure 5.8]{tm-0728}.}
  \label{0728-fig5-8}
\end{figure}

>>>>

\documentclass[ms,a4paper]{memoir}
\chapterstyle{dash}
\usepackage{ulem}
\usepackage{xcolor}
\usepackage[a4paper]{geometry}
\usepackage{float}

\renewcommand{\thesection}{\arabic{section}}
\maxtocdepth{subsection}
\counterwithout{figure}{chapter}
\counterwithout{table}{chapter}

@@

\title{msre reactor design}
\author{openmsr}
\date{}
\pagestyle{empty}
\addtocontents{toc}{\protect\thispagestyle{empty}}

@@

\vspace{-4cm}
\renewcommand{\contentsname}{contents}
\tableofcontents*

\section{\emph{introduction}}
the design and dimensions of the molten salt reactor experiment is detailed in several ornl reports. this document aims at providing a comprehensive overview of the core design, dimensions, and materials, for the purpose of creating an accurate cad model of the reactor core. \\

@@

\section{\emph{overview}}
an isometric view of the msre core can be seen in figure \ref{3229-fig1}
\begin{figure}[H]
  \centering
  \fbox{\includegraphics[page=6,width=0.8\textwidth,trim={2cm 6cm 3.5cm 4cm},clip]{\mrsarchive/docs/ORNL-TM-3229.pdf}}
  \caption{msre reactor vessel \parencite[figure 1]{ornl-tm-3229}.}
  \label{3229-fig1}
\end{figure}

@@

\section{\emph{core vessel}}
\subsection{\emph{lower and upper vessel plenum}}
\textcite[page 14]{ornl-tm-3229} states about the lower vessel plenum:
\enquote{\textit{the lower plenum of the core vessel is formed by a standard 60 in. od asme flanged and dished head}} and \textcite[page 25]{ornl-tm-3229} states about the upper vessel plenum:
\enquote{\textit{the reactor vessel upper plenum is similar to the lower plenum in that it is forme from a standard asme flanged and dished head}}.

\subsection{\emph{flow distributor and annulus}}
\textcite[page 112 and 114]{ornl-tm-3039} states about the annulus:
\enquote{\textit{the fuel salt entered the flow distributor where it was directed downward around the circumference of the vessel. it flowed in a spiral path through a 1-in. annulus between the vessel wall and the core can for cooling purposes.}}. \enquote{\textit{to prevent possible overheating in an otherwise stagnant region, a small portion of salt entering the reactor was diverted into the region just above the core can support flange in the annulus between the vessel and the core can}}.
this is depicted in figure \ref{0728-p109-p110}. \\

@@

\subsubsection{\emph{flow distributor orifices}}
\textcite[page 9-10]{ornl-tm-3229} states about the flow distributor orifices:
\enquote{\textit{one characteristic of this type of volute is the variable static pressure around it; therefore, in order to obtain a uniform angular flow distribution, it is necessary to use orifices between the volute and core wall cooling annulus. the orifices are 3/4 in. and occur in stacks of 3. at the head end of the volute the orifice stacks are 5 deg apart. at the tail end of the volute, because of the lower fuel velocity and resulting higher static pressure, the orifice stack spacing is increased to 22 1/2 deg. this orifice distribution was determined from the 1/5 scale model. the orifice holes are drilled at an angle of 30 deg with the tangent in order to maintain a tangential velocity component in the annulus. the resulting high heat transfer coeffficients cool the core vessel wall and the reactor core can}}.

@@

\subsubsection{\emph{bypass flow, core support lugs and flange}}
\parencite[page 85]{ornl-tm-0728} states about the bypass:
\enquote{\textit{to prevent possible overheating in a region that might otherwise have been stagnant, about 24 gpm of salt entering the reactor is diverted into the region just above the core-can support flange in the annulus between the pressure vessel and the core can. this is accomplished through 18 slots or channels, 0.2 by 0.2 in., cut in the core-can flange. these slots are machined at an angle of $30^\circ$ to promote better mixing in the region. in addition, a by-pass flow of 3-22 gpm of salt will pass through the annular clearances at the core can support ring}}.

@@

\subsection{\emph{anti-swirl vanes}}
\textcite[page 14]{ornl-tm-3229} states about the anti-swirl vanes:
\enquote{\textit{the anti-swirl vanes consist of 48 plates starting about 2 in. up in the core wall cooling annulus and extending along radial lines into the lower head for about 38\% of the radial distance to the core centerline. they are slightly elevated of the core vessel wall, thus eliminating as much coner area as possible where settled solids could accumulate.}}. \\

@@

\subsection{\emph{vessel drain}}
\label{sec:drain}
\textcite[page 14-15]{ornl-tm-3229} states about the the two drain lines:
\enquote{\textit{the vessel drain consist of a short section of 1 1/2 in. pipe extending slightly up into the vessel head at the centerline, and having a conical umbrella over it. in addition it incorporates a secondary drain with is a 1/2 in. contentric tube coming up the middel of the primary drain, penetrating through the drain pipe wall just inside the vessel and wrapping around the drain pipe horizontally about $90^\circ$}}.

@@

\subsection{\emph{grid support plates}}
\parencite[page 20]{ornl-tm-3229} states about the grid support structure:
\enquote{\textit{in flowing through the reactor moderator assembly, fuel first passes through a moderator support structure and then through the moderator core blocks. the support structure consists of two assemblies, the lower of which is a hastelloy-n cross structure ... which is the main support structure for the graphite}}.

@@

\begin{figure}[H]
  \centering
  \fbox{\includegraphics[page=165,width=0.8\textwidth,trim={4cm 5.5cm 4cm 2.25cm},clip]{\mrsarchive/docs/ORNL-4676.pdf}}
  \caption{comparison of the bottom positioning stud on the graphite moderator element when initially installed in the msre and after operation \parencite[figure 13.3]{ornl-4676}.}
  \label{4676-fig13-3}
\end{figure}

@@

\subsection{\emph{vessel outlet}}
\parencite[page 84-85]{ornl-tm-0728} states about the vessel outlet:
\enquote{\textit{the salt leaves the reactor core and flows through the upper head to the 10-in. nozzle opening. it is diverted through a 5-in. opening in the side of the nozzle to flow to the fuel circulating pump. the 10-in. nozzle also serves as an access port and support for the tree control rods and for taking and placing of the four graphite-sample rods in the core matrix}}.

@@

\begin{figure}[H]
  \centering
  \fbox{\includegraphics[page=145,width=0.8\textwidth,trim={4cm 5cm 4cm 3cm},clip]{\mrsarchive/docs/ORNL-TM-0728.pdf}}
  \caption{control rod and drive assembly \parencite[figure 5.14]{ornl-tm-0728}.}
  % \label{3708-fig40}
\end{figure}

% figure \ref{0728-p111-p112} shows a section view of the msre core
\begin{figure}[H]
  \centering
  \fbox{
  \begin{tabular}{c}
    \includegraphics[page=111,width=0.8075\textwidth,trim={2.5cm 0.04cm 2.5cm 10cm},clip]{\mrsarchive/docs/ORNL-TM-0728.pdf}\\[-0.12cm]
    \includegraphics[page=112,width=0.8\textwidth,trim={2.64cm 31cm 2.5cm 0.04cm},clip]{\mrsarchive/docs/ORNL-TM-0728.pdf}
  \end{tabular}
  }
  \caption{elevation of cotnrol rod drive housings \parencite[page 111-112]{ornl-tm-0728}.}
  \label{0728-p111-p112}
\end{figure}

\begin{figure}[H]
  \centering
  \fbox{\includegraphics[page=87,width=0.8\textwidth,trim={3cm 4cm 3.75cm 3cm},clip]{\mrsarchive/docs/ORNL-3708.pdf}}
  \caption{reactor completed and ready for shipment to msre site \parencite[figure 44]{ornl-3708}.}
  % \label{3708-fig40}
\end{figure}

@@

\subsubsection{\emph{removable nozzle plug}}
\parencite[page 104-105]{ornl-tm-0728} states about the removable nozzle plug:
\enquote{\textit{the removable nozzle plug is about 20-1/2 in. long and is 9.770 in od at the top and 9.520 in. od at the bottom to provide a taper to assist in freeing the plug for removal. the radial clearance is 1/8 in. at the top and 1/4 in. at the bottom. fuel salt is frozen in the annular space between the plug and the nozzle, to effect a salt seal. the salt maintained below the freezing temperature by cooling the outside of the nozzle extension and the inside of the plug with a flow of cell atmosphere gas}}.
\enquote{\textit{the salt seal prevents the salt from coming in contact with and corroding the ring-joint gas seal on the upper flange}}.
\enquote{\textit{the nozzle plug is hollow and is filled with insulation. the upper end of the plug is welded to the mating flange for the 10-in. closure}}.

@@

\subsubsection{\emph{outlet strainer}}
\textcite[page 24-26]{ornl-tm-3229} states about the outlet strainer:
\enquote{\textit{after the model was built, a strainer structure ... was added to the reactor outlet design which penetrated down into the upper plenum. its purpose was to catch graphite chips (graphite floats in fuel salt) should they break loose from the moderator. this device was not installed in the model but it would not be expected to incluence the flow significantly}}. \\

@@

\textcite[page 115]{ornl-tm-3039} states about the outlet strainer:
\enquote{\textit{stainers were provided at the reactor outlet to prevent passage of graphite chips larger then 1/16 in. a strainer basket was attached to the lower end of the 10-in. nozzle plug and extended downward into the upper head region of the reactor vessel. the three control rod thimbles and the graphite sample assembly passed through the stainer basket. since the five graphite elements were not pinned as were the remaining moderator elements, a cross-shaped extension of the basket essembly projected beneath the basket to provide a hold-down}}. this is depicted in figure \ref{3039-fig5-11}. \\

@@

\section{\emph{graphite can}}
the msre used a graphite moderator consisting of an assembly of graphite stringer with control rods and removable samples rods in the center. bellow is seen the assembly of the can without the removable rods at the center.
\begin{figure}[H]
  \centering
  \fbox{
    \includegraphics[width=0.4\textwidth]{\mrsarchive/img/core-half.jpg}
    \includegraphics[width=0.393\textwidth]{\mrsarchive/img/core-full.jpg}
  }
  \caption{msre graphite core assembly \parencite{ornl}.}
  \label{core-assembly}
\end{figure}
\textcite[page 112 and 114]{ornl-tm-3039} states about the graphite stringers: \enquote{\textit{the reactor core consisted of 617 full and factional sized graphite elements 2x2 in. cross section and about 67 in. long}}. \enquote{\textit{when not buoyed by being immersed in salt, the vertical graphite elements were supported by a lattice of graphite blocks which in turn are supported by a hastelloy-n grid fastened to the bottom of the core can. the core can was thus free to expand downward relative to its top attachment to the reactor vessel while the graphite was free to expand upward relative to its support at the lower end of the core can. the graphite moderator elements were restrained from floating out of core by hastelloy-n rod through holes in the dowel section at the bottom of each graphite element. in addition a wire passing through the upper graphite elements prevented the upper partion of a broken element from floating out of the core}}.

@@

\subsection{\emph{typical graphite stringer}}
a typical graphite stringers, along with a removable graphite stringer, is given in figure \ref{3229-fig2}
\begin{figure}[H]
  \centering
  \centering
  \fbox{\includegraphics[page=7,width=0.8\textwidth,trim={3.5cm 6cm 2.5cm 5cm},clip]{\mrsarchive/docs/ORNL-TM-3229.pdf}}
  \caption{typical graphite core block arrangement \parencite[figure 2]{ornl-tm-3229}.}
  \label{3229-fig2}
\end{figure}

@@

\subsubsection{\emph{dowel section}}
the dowel at the bottom consists of two section dowel (long slim part and stumpy thick part) seen below.
\begin{figure}[H]
  \centering
  \fbox{\includegraphics[page=385,width=0.8\textwidth,trim={2.5cm 6.2cm 3.3cm 4.1cm},clip]{\mrsarchive/docs/ORNL-3708.pdf}}
  \caption{appearance of spalls and cracks in samples of severly cracked grade cgb graphite in (a) spindle end of a machined bar and (b) midplane of an as-received bar. cracks are indicated by arrows or are clearly visible as ragged white lines. \parencite[figure 4]{ornl-3708}.}
  \label{3708-fig4}
\end{figure}

@@

\subsection{\emph{hold-down rods and support wires}}
\label{sec:hold-down-rods-support-wires}
\parencite[page 84-85]{ornl-tm-0728} states about the hold-down rods and support wires:
\enquote{\textit{a 5/16-in.-diam inor-8 rod passes through a 0.010-in.-wall-thichness bushing placed in a 0.375-in.-diam hole in the dowel section at the bottom of each graphite stringer. these rods also pass through the inor-8 grid-supporting structure and prevent each graphite stringer from floating in the fuel salt. if a graphite stringer were to break in two, the top partion would tend to float away and leave a relatively stagnant pocket of fuel salt which might reach a higher temperature than desired. the effect of this on the reactivity and on the temperatures in the reactor has been studied. to guard against this eventuality, a 1/16-in.-diam inor-8 wire is passed through a 1/8-in.-diam inor-8 insert about 1 in from the top of each graphite stringer, fastening the tops together and to the core can}}.

@@

\subsection{\emph{graphite lattice blocks}}
\parencite[page 20]{ornl-tm-3229} states about the lattice blocks: \enquote{\textit{resting on the this}(grid support plates)\textit{ is a grid consisting of two layers of rectangular graphite bars, one layer resting on the other and perpendicular to it. the purpose of the graphite assembly is to position and hold the graphite core blocks, and to compensate for a difference in thermal expansion between the hastelloy n and graphite}}.

@@

\subsection{\emph{removable graphite stringer}}
\parencite[page 84]{ornl-tm-0728} states about the removable graphite stringer:
\enquote{\textit{the five graphite stringers at the center of the core can be removed, although with considerable more difficulty}}.
see figure \ref{0728-fig5-7}.
\enquote{\textit{the five stringers are of a special design (types 7, 60 and 61 on ornl dwgs d-bb-b-40416, 40418 and 40581). they are 2-in. x 2-in. in cross section but are 64-1/2 in. long rather than the 62-1/8 in. of the average stringer. they do not have the dowel section at the bottom and the hole for the hold-down rods and they rest directly on the inor-8 supporting grid rather than on the graphite lattice blocks. the lattice blocks do not extend across the five stringers locations, the opening thus provided through the blocks permitting insertion of a viewing device through the core to permit observation of the lower head should this prove desirable}}.
\enquote{\textit{each of the five stringers is drilled and tapped with a 3/4-in. 6 acme thread on the upper end for an inor-8 lifting knob, or stud, which can engage a snaptite quick-disconnect coupling. they are prevented from floating in the fuel salt by 1/8 in. x 1/2 in. hold-down bars welded to the strainer assembly at a level even with the top of the graphite core. (see ornl dwg e-bb-b-40598)}}.
see the figure \ref{3229-fig2}.

@@

\begin{figure}[H]
  \centering
  \centering
  \fbox{\includegraphics[page=164,width=0.8\textwidth,trim={3.5cm 11.5cm 2cm 3.5cm},clip]{\mrsarchive/docs/ORNL-4676.pdf}}
  \caption{photograph of the top of the graphite moderator element, showing the crack originating at the INOR-8 lifting stud and extending partway around the graphite block shown by the short arrows. The long arrow indicates the same channel in the different views \parencite[figure 13.2]{ornl-4676}.}
  \label{4676-fig13-2}
\end{figure}

% \subsection{\emph{tall graphite stringers}

\subsection{\emph{core centering grid or graphite centering bridge}}
\parencite[page 84-85]{ornl-tm-0728} states about the core centering grid:
\enquote{\textit{at the top of the graphite blocks a centering bridge holds a row of stringers in position on two diameters at right angles to each other. this bridge helps to prevent shifting of the entire stringer assembly (see ornl dwg. d-bb-b-40424)}}.
\begin{figure}[H]
  \centering
  \centering
  \fbox{\includegraphics[page=29,width=0.8\textwidth,trim={3cm 13.7cm 4cm 4cm},clip]{\mrsarchive/docs/ORNL-4676.pdf}}
  \caption{fish eye view of top of core assembly with centering bridge and broken pieces of graphite \parencite[figure 1.5]{ornl-4676}.}
  \label{4676-fig1-5}
\end{figure}

@@

\subsection{\emph{restrictor ring}}
\parencite[page 84-85]{ornl-tm-0728} states about the restrictor ring:
\enquote{\textit{to prevent an excessive amount of salt flow in the annuluar space thus created, an inor-8 restrictor ring, 1/2 x 1/2 x 54-1/2 in. id. surrounds the bundle of graphite stringers at the bottom (see ornl dwg. d-bb-b-40427). the stringers are restrained from excessive movement at the top by a graphite retainer ring, 1 x 2 x 53-1/4 in. id. this ring, in turn, is held in place by an inor-8 retainer ring, 3/4 x 3/4 x 53-1/4 in. id (see ornl dwg. d-bb-b-40428 for both rings)}}.

\subsection{\emph{core can}}
\parencite[page 79]{ornl-tm-0728} states about the core can:
\enquote{\textit{the core can, or shell, is 55-1/2 in. ID and 67-15/16 in. high and was rolled from 1/4-in.-thick INOR-8 plate.}}

\section{\emph{control stringer and sample assembly}}
the control stringer and sample assembly is further detailed in figure \ref{61246-fig14a}, \ref{61246-fig14b}, and \ref{61246-fig14c}.

figure \ref{61246-fig14a} depicts what seems to be a early design of the stringer arrangement
\begin{figure}[H]
  \centering
  \fbox{\includegraphics[page=23,width=0.8\textwidth,trim={4cm 2cm 4.8cm 2cm},clip]{\mrsarchive/docs/ORNL-CF-61-2-46-AD.pdf}}
  \caption{msre control stringer arrangement \parencite[figure 14a]{ad-cf-61-2-46}.}
  \label{61246-fig14a}
\end{figure}
later in 1965 figure \ref{0728-fig5-7} give the stringer arrangement
\begin{figure}[H]
  \centering
  \fbox{\includegraphics[page=117,width=0.8\textwidth,trim={2.7cm 5.5cm 3.1cm 4.2cm},clip,angle=0.4]{\mrsarchive/docs/ORNL-TM-0728.pdf}}
  \caption{lattice arrangement of control rods and sample rod \parencite[figure 5.7]{ornl-tm-0728}.}
  \label{0728-fig5-7}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \fbox{
    \begin{tabular}{c}
      \includegraphics[page=118,width=0.8\textwidth,trim={11cm 5cm 0cm 4cm},clip,angle=-90]{\mrsarchive/docs/ORNL-TM-0728.pdf} \\ [-0.3cm]
      \includegraphics[page=119,height=0.50\textwidth,trim={0cm 4.4cm 33cm 3cm},clip,angle=-90]{\mrsarchive/docs/ORNL-TM-0728.pdf}
    \end{tabular}
  }
  \caption{graphite-inor-8 sample assembly \parencite[figure 5.8]{ornl-tm-0728}.}
  \label{0728-fig5-8}
\end{figure}

% \begin{figure}[H]
%   \centering
%   \centering
%   \fbox{\includegraphics[page=130,width=0.8\textwidth,trim={3.8cm 3.7cm 3.5cm 2.5cm},clip,angle=0.25]{\mrsarchive/docs/ORNL-TM-3039.pdf}}
%   \caption{earlier design of msre surveillance facility inside reactor vessel \parencite[figure 5.14]{ornl-tm-3229}.}
%   \label{3039-fig5-11}
% \end{figure}

\begin{figure}[H]
  \centering
  \centering
  \fbox{\includegraphics[page=107,width=0.8\textwidth,trim={4cm 22cm 1.5cm 5cm},clip,angle=0.25]{\mrsarchive/docs/ORNL-3872.pdf}}
  \caption{three stringers (s, r, and l) of cgb graphite and inor-8 surveillance specimens \parencite[figure 4.7]{ornl-3872}.}
  % \label{3039-fig5-11}
\end{figure}
